Pathogen Spillover - dissertation chapter

### 2016-09-09 pathogen-spillover-diss-ch.Rmd
Add figure printing for temperature and rainfall.

@@

```{r hours >25 path model no umca density scaled}
model_list2.scl <- list(
      shannons <- lm(shannons2005.scl ~ twi15m.log.scl, data = oakplots.sub),

@@

**Residual plots of symptomatic leaf count model**
```{r symptomatic leaf count residuals plot, echo=F}
plot(slc, main = "LMM of symptomatic leaf count residuals")
plot(slc@frame$umca_density_ha.log.scl, resid(slc),
     xlab = "bay laurel density",
     ylab = "model residuals",
     main = "SLC model residuals vs. bay laurel density predictor")
plot(slc@frame$hrs_abv25ds.scl, resid(slc),
     xlab = "# dry-season hours >25 C",
     ylab = "model residuals",
     main = "SLC model residuals vs. dry-season hours >25 C predictor")
plot(slc@frame$avg_hrs1422_wet_t_t1.log.scl,
     xlab = "avg. wet-hours 14-22C",
     ylab = "model residuals",
     main = "SLC model residuals vs. wet-hours predictor")
plot(slc@frame$shannons2005.scl, resid(slc),
     xlab = "Shannon's diversity index",
     ylab = "model residuals",
     main = "SLC model residuals vs. Shannon's diversity predictor")
plot(slc@frame$twi15m.log.scl, resid(slc),
     xlab = "wetness index",
     ylab = "model residuals",
     main = "SLC model residuals vs. wetness index predictor")
```
```{r pairs plot of slc predictor variables}
pairs(select(oakplots.sub, umca_density_ha.log.scl, hrs_abv25ds.scl, avg_hrs1422_wet_t_t1.log.scl, shannons2005.scl, twi15m.log.scl),
      lower.panel = panel.cor, diag.panel = panel.hist,
      main = "Predictor variables for symptomatic leaf count models")
```

@@

**Residual plots of oak infection model**
```{r oak infection residuals plot, echo=F}
plot(oak_infection, main = "Binomial GLMM of oak infection residuals")
plot(oak_infection@frame$tot_lfct.log.scl, resid(oak_infection),
     xlab = "symptomatic leaf count",
     ylab = "model residuals",
     main = "Oak infection model residuals vs. symptomatic leaf count predictor")
plot(oak_infection@frame$avg_hrs1422_wet_t_t1.log.scl, resid(oak_infection),
     xlab = "avg. wet-hours 14-22C",
     ylab = "model residuals",
     main = "Oak infection model residuals vs. wet-hours predictor")
plot(oak_infection@frame$shannons2005.scl, resid(oak_infection),
     xlab = "Shannon's diversity index",
     ylab = "model residuals",
     main = "Oak infection model residuals vs. diversity predictor")
```
```{r pairs plot of oak infection predictor variables, echo=F}
pairs(select(oakplots.sub, tot_lfct.log.scl, avg_hrs1422_wet_t_t1.log.scl, shannons2005.scl),
      lower.panel = panel.cor, diag.panel = panel.hist,
      main = "Predictor variables for oak infection model")
```

@@

```{r fit and compare all oak plots, echo=F, message=F}
raw_variables_model <- sem.fit(model_list1, oakplots.sub, 
        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
scaled_variables_model <- sem.fit(model_list1.scl, oakplots.sub,
          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))

raw_variables_model_noUMCA <- sem.fit(model_list2, oakplots.sub, 
        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
scaled_variables_model_noUMCA <- sem.fit(model_list2.scl, oakplots.sub,
          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))
```

```{r tables of d-sep test results from sem.fit, echo=F}
kable(raw_variables_model, 
      caption = "Results of the d-sep test using raw variables.")
kable(scaled_variables_model,
      caption = "Results of the d-sep test using standardized variables.")
```

@@

```{r get path coefficients, echo=F}
raw_coefficients <- sem.coefs(model_list1, oakplots.sub,
          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
scaled_coefficients <- sem.coefs(model_list1.scl, oakplots.sub,
          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))
```

@@

```{r tables of path coefficients, echo=F}
library(knitr)
kable(raw_coefficients,
      caption = "Raw path coefficients")
kable(scaled_coefficients,
      caption = "Standardized path coefficients")
```

@@

## Temperature Figures
```{r temperature figures, echo=F}
library(ggplot2)
library(scales)
temperatures <- left_join(
      select(summer_temps, plotid, sample_year, hrs_abv25ds),
      select(wet_hrs, plotid, sample_year, hrs1422_wet),
      by = c("plotid","sample_year"))
# oak_sod$year <- as.Date(oak_sod$sample_year)

@@

sstemps_figure <- temps_figure + 
      geom_boxplot(aes(as.factor(sample_year), hrs_abv25ds)) +
      theme_bw() +
      xlab("Sample Year") +
      ylab("Number of hours") +
      ggtitle("Dry-season Hours >25 C")

slc_avg_figure <- temps_figure + 
      geom_boxplot(aes(as.factor(sample_year), avg_lfct)) +
      theme_bw() +
      xlab("Sample Year") +
      ylab("Symptomatic leaf count") +
      ggtitle("California bay laurel leaf symptoms - plot averages")

slc_total_figure <- temps_figure + 
      geom_boxplot(aes(as.factor(sample_year), tot_lfct)) +
      theme_bw() +
      xlab("Sample Year") +
      ylab("Symptomatic leaf count") +
      ggtitle("California bay laurel leaf symptoms - plot totals")

rstemps_figure <- temps_figure + 
      geom_boxplot(aes(as.factor(sample_year), hrs1422_wet)) +
      theme_bw() +
      xlab("Sample Year") +
      ylab("Number of hours") +
      ggtitle("Wet-hours 14-22 C")

oak_infection_figure <- temps_figure +
      geom_bar(aes(as.factor(sample_year), inf_oak_ct), stat = "identity") +
      # geom_boxplot(aes(as.factor(sample_year), inf_oak_ct)) +
      theme_bw() +
      xlab("Sample Year") +
      ylab("Number of infected oaks") +
      ggtitle("Incident oak infection")
```

## Rainfall Figures
```{r rainfall figures}
# summary(rainy_season_ppt)
```

```{r print session info}
print(sessionInfo())
```

>>>>

panel.cor <- function(x, y, digits = 2, cex.cor, ...)
{
      usr <- par("usr"); on.exit(par(usr))
      par(usr = c(0, 1, 0, 1))
      # correlation coefficient
      r <- cor(x, y)
      txt <- format(c(r, 0.123456789), digits = digits)[1]
      txt <- paste("r= ", txt, sep = "")
      text(0.5, 0.6, txt)

      # p-value calculation
      # p <- cor.test(x, y)$p.value
      # txt2 <- format(c(p, 0.123456789), digits = digits)[1]
      # txt2 <- paste("p= ", txt2, sep = "")
      # if(p<0.01) txt2 <- paste("p= ", "<0.01", sep = "")
      # text(0.5, 0.4, txt2)
}

@@

```{r hours >25 path model no umca density scaled, echo=F}
model_list2.scl <- list(
      shannons <- lm(shannons2005.scl ~ twi15m.log.scl, data = oakplots.sub),

@@

**Residual plots of symptomatic leaf count model**
```{r symptomatic leaf count residuals plot, echo=F}
plot(resid(slc), main = "LMM of symptomatic leaf count residuals")
# qqnorm(resid(slc), main="Q-Q plot for conditional residuals - SLC")
# qqnorm(ranef(slc)$sample_year$`(Intercept)`, 
#        main="Q-Q plot for Sample Year")
# qqnorm(ranef(slc)$plotid$`(Intercept)`, 
#        main="Q-Q plot for plot ID")

plot(slc@frame$umca_density_ha.log.scl, resid(slc),
     xlab = "bay laurel density",
     ylab = "model residuals",
     main = "SLC model residuals vs. bay laurel density predictor")
plot(slc@frame$hrs_abv25ds.scl, resid(slc),
     xlab = "# dry-season hours >25 C",
     ylab = "model residuals",
     main = "SLC model residuals vs. dry-season hours >25 C predictor")
plot(slc@frame$avg_hrs1422_wet_t_t1.log.scl,
     xlab = "avg. wet-hours 14-22C",
     ylab = "model residuals",
     main = "SLC model residuals vs. wet-hours predictor")
plot(slc@frame$shannons2005.scl, resid(slc),
     xlab = "Shannon's diversity index",
     ylab = "model residuals",
     main = "SLC model residuals vs. Shannon's diversity predictor")
plot(slc@frame$twi15m.log.scl, resid(slc),
     xlab = "wetness index",
     ylab = "model residuals",
     main = "SLC model residuals vs. wetness index predictor")
```
```{r pairs plot of slc predictor variables, echo=F}
pairs(na.omit(select(oakplots.sub, umca_density_ha.log.scl, hrs_abv25ds.scl, avg_hrs1422_wet_t_t1.log.scl, shannons2005.scl, twi15m.log.scl)),
      lower.panel = panel.cor, diag.panel = panel.hist,
      main = "Predictor variables for symptomatic leaf count models")
```

@@

**Residual plots of oak infection model**
```{r oak infection residuals plot, echo=F}
plot(oak_infection, main = "Binomial GLMM of oak infection residuals")
plot(oak_infection@frame$tot_lfct.log.scl, resid(oak_infection),
     xlab = "symptomatic leaf count",
     ylab = "model residuals",
     main = "Oak infection model residuals vs. symptomatic leaf count predictor")
plot(oak_infection@frame$avg_hrs1422_wet_t_t1.log.scl, resid(oak_infection),
     xlab = "avg. wet-hours 14-22C",
     ylab = "model residuals",
     main = "Oak infection model residuals vs. wet-hours predictor")
plot(oak_infection@frame$shannons2005.scl, resid(oak_infection),
     xlab = "Shannon's diversity index",
     ylab = "model residuals",
     main = "Oak infection model residuals vs. diversity predictor")
```
```{r pairs plot of oak infection predictor variables, echo=F}
pairs(na.omit(select(oakplots.sub, tot_lfct.log.scl, avg_hrs1422_wet_t_t1.log.scl, shannons2005.scl)),
      lower.panel = panel.cor, diag.panel = panel.hist,
      main = "Predictor variables for oak infection model")
```

@@

```{r fit and compare all oak plots, echo=F, message=F, warning=F}
raw_variables_model <- sem.fit(model_list1, oakplots.sub, 
        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
        .progressBar = F)
scaled_variables_model <- sem.fit(model_list1.scl, oakplots.sub,
          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"),
          .progressBar = F)

raw_variables_model_noUMCA <- sem.fit(model_list2, oakplots.sub, 
        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
        .progressBar = F)
scaled_variables_model_noUMCA <- sem.fit(model_list2.scl, oakplots.sub,
          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"),
          .progressBar = F)
```

```{r tables of d-sep test results from sem.fit, echo=F}
kable(raw_variables_model, 
      caption = "Results of the d-sep test using raw variables.")
kable(scaled_variables_model,
      caption = "Results of the d-sep test using standardized variables.")

kable(raw_variables_model_noUMCA, 
      caption = "Results of the d-sep test using raw variables - no UMCA density.")
kable(scaled_variables_model_noUMCA,
      caption = "Results of the d-sep test using standardized variables - no UMCA density.")
```

@@

```{r get path coefficients, echo=F}
raw_coefficients <- sem.coefs(model_list1, oakplots.sub,
          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
scaled_coefficients <- sem.coefs(model_list1.scl, oakplots.sub,
          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))

raw_coefficients_noUMCA <- sem.coefs(model_list2, oakplots.sub,
          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
scaled_coefficients_noUMCA <- sem.coefs(model_list2.scl, oakplots.sub,
          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))
```

@@

```{r tables of path coefficients, echo=F}
kable(raw_coefficients,
      caption = "Raw path coefficients")
kable(raw_coefficients_noUMCA,
      caption = "Raw path coefficients - no UMCA density")

kable(scaled_coefficients,
      caption = "Standardized path coefficients")
kable(scaled_coefficients_noUMCA,
      caption = "Standardized path coefficients - no UMCA density")
```

@@

## Temperature Figures
```{r temperature figures, echo=F, warning=F, eval=F}
library(ggplot2)
library(scales)
temperatures <- left_join(
      select(summer_temps, plotid, sample_year, hrs_abv25ds),
      select(wet_hrs, plotid, sample_year, hrs1422_wet),
      by = c("plotid","sample_year"))
# oak_sod$year <- as.Date(oak_sod$sample_year)

@@

(sstemps_figure <- temps_figure + 
      geom_boxplot(aes(as.factor(sample_year), hrs_abv25ds)) +
      theme_bw() +
      xlab("Sample Year") +
      ylab("Number of hours") +
      ggtitle("Dry-season Hours >25 C")
)     
(slc_avg_figure <- temps_figure + 
      geom_boxplot(aes(as.factor(sample_year), avg_lfct)) +
      theme_bw() +
      xlab("Sample Year") +
      ylab("Symptomatic leaf count") +
      ggtitle("California bay laurel leaf symptoms - plot averages")
)
(slc_total_figure <- temps_figure + 
      geom_boxplot(aes(as.factor(sample_year), tot_lfct)) +
      theme_bw() +
      xlab("Sample Year") +
      ylab("Symptomatic leaf count") +
      ggtitle("California bay laurel leaf symptoms - plot totals")
)
(rstemps_figure <- temps_figure + 
      geom_boxplot(aes(as.factor(sample_year), hrs1422_wet)) +
      theme_bw() +
      xlab("Sample Year") +
      ylab("Number of hours") +
      ggtitle("Wet-hours 14-22 C")
)
(oak_infection_figure <- temps_figure +
      geom_bar(aes(as.factor(sample_year), inf_oak_ct), stat = "identity") +
      # geom_boxplot(aes(as.factor(sample_year), inf_oak_ct)) +
      theme_bw() +
      xlab("Sample Year") +
      ylab("Number of infected oaks") +
      ggtitle("Incident oak infection")
)
```

## Rainfall Figure
Total annual rainfall estimated at plot locations using two-dimensional interpolation methods implemented in GRASS-GIS. The red line indicates the average rainfall from 2005-2014
```{r plot rainfall figure, echo=F}
library(ggplot2)
# summary(rainy_season_ppt)
# rainy_season_ppt$year <- as.factor(rainy_season_ppt$year)
rainfall_figure <- ggplot(filter(rainy_season_ppt, year > 2004), 
                          aes(year, rain2d_total))
rainfall_figure +
      geom_boxplot(aes(group = year)) +
      geom_hline(aes(yintercept = mean(rain2d_total)),
                 colour = "red") +
      xlab("Sample Year") +
      ylab("November-May Total Rainfall (cm)") +
      ggtitle("Rainfall Estimated at Plots from Rain Gauges")
```

### Session Information
```{r print session info, echo=FALSE}
print(sessionInfo())
```
